### 2017-02-14 hoare/paper.tex
Init after deadline

\documentclass[twoside,twocolumn]{article}

\usepackage{blindtext} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

% \usepackage{fancyhdr} % Headers and footers
% \pagestyle{fancy} % All pages have headers and footers
% \fancyhead{} % Blank out the default header
% \fancyfoot{} % Blank out the default footer
% \fancyhead[C]{} % Custom header text
% \fancyfoot[C]{\thepage} % Custom footer text

\usepackage{dirtytalk}
\renewcommand{\quote}[2]{
  \say{
    {\itshape #1}
  }\\[.5em]
  \rightline{{\rm --- #2}}

  \vspace{1em}
}

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Large\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{An Axiomatic Basis for Computer Programming: Review} % Article title
\author{%
\textsc{Gosha Krikun}\\[1ex] % Your name
\normalsize Innopolis University\\ % Your institution
\normalsize \href{mailto:g.krikun@innopolis.ru}{g.krikun@innopolis.ru} % Your email address
%\and % Uncomment if 2 authors are required, duplicate these 4 lines if more
%\textsc{Jane Smith}\thanks{Corresponding author} \\[1ex] % Second author's name
%\normalsize University of Utah \\ % Second author's institution
%\normalsize \href{mailto:jane@smith.com}{jane@smith.com} % Second author's email address
}
% \date{February 10, 2017} % Leave empty to omit a date
\date{\today}
\renewcommand{\maketitlehookd}{%
\begin{abstract}
  \noindent

  This is review on C.A.R. Hoare's paper, An Axiomatic Basis for Computer
  Programming.

\end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

% Print the title
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}

``People are not perfect'', this is what we learned so fourth. 
Every one could make a mistake. In some areas it could be unnoticed, but in
others it could cause serious consequences.

There are a lot of people with their needs. Majority of them lie outside
safety-critical sector, error and failures in which could leads to deaths. 
People won't spent big money for time of qualified developers to avoid minor
bugs.

So nowadays main approach to find failures and errors is testing.
And I can't not mentioned my favorite quote about it:\\

\quote{
Program testing can be used to show the presence of bugs, but never to show their absence!
}{Edsger Wybe Dijkstra}

But more and more critical tasks entrust to computer systems. When errors and
failures cost increases, we will pay for diligence and accuracy as much as
needed.

To be exactly sure, that we don't make mistakes we make premises and develop techniques for
reasoning about properties and correctness. And for proving it we used our logic
and deduction.

\section{Purpose and Applications}

So Hoare suggested triple \{P\} Q \{R\} \cite{hoare:69} in which developer need
define 
precondition (premises) and postcondition (conclusion) which hold for program Q
after execution.

This technique helps prove correctness of a program Q, according to defined
logical predicates P and R.

It doesn't help derive true predicates from requirements, but it
provide ability to use common proof techniques for deducting correctness.

For next disscussion suppose that we define right P and R from some
requirements. \\

Speaking about correctness we should understand, that any answer could be in
three state: answer received and it is correct, answer not received, answer is
incorrect.

So when answer received and it is right, for all input values in the domain,
then we call it total correctness.

If answer could be not received (e.g. program doesn't terminate) for some values
in the domain, but is right for others - partial correctness.

And incorrectness in last case. \\

Using axioms and inference rules what was proposed by Hoare, we could deduct
that if P holds before execution of Q, and if Q terminates, then R will holds
after execution (correctness of Q).\\

In 1969 paper Hoare didn't mentioned that Q should terminates, so he defined
rules just for deducting partial correctness (particular in rule D3). But later
he add definition for total correctness too. \\

For example lets take task on finding real roots for quadratic equation.

We limits input set according to discriminant and set of real numbers in
predicate P. Then we could find values by some program Q. And finally check
whether equation holds or not in R.

\section{Axioms and Rules}

In the article Hoare states four basic rules:
\begin{itemize}
\item
  Axiom of Assignment
\item
  Rule of Consequence
\item
  Rule of Composition
\item
  Rule of Iteration
\end{itemize}

Also it requires no side effects for Q (nothing changes except state which
we define in pre and postconditions). Because in other case, we couldn't
consider correctness of overall state. \\

Using this rules we can make deductive proof from our premise to conclusion,
that R will holds after execution.

But only after, not during.

It could be a case when program Q during execution
change state which doesn't hold conditions P and R.

This case could appear when specification describe something in general, but
doesn't divide execution into small pieces of actions.

If we think about it in more details, we could decompose it to two sequence of
statements (rule of composition D2), in which first part hold some postcondition
R1 and second have R1 as precondition.

So nothing scary in that case if we still could prove correctness of our program.

P like a guard, filter input on what program will work correctly, and R check
results to be correct.

% But sometimes it could be not so straightforward. Lets return to our example
% about real roots. Suppose quadratic coefficient a could be equals 0.

% In that case equation is linear but still valid, there is solution when equation holds.
% But if we solve it by general formula (formula)

% Rules.
% There is set of rules (or axioms) which we could use to deduct correctness of a program.

However precondition and postconditions could be almost same. This looks ugly.

firstly we should define the
truth.

So directed tests could show that implementation isn't work properly. This is
helpful during development and evolving software product.

%------------------------------------------------

\section{Shortcomings}

\section{Conclusion}

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Hoa69]{hoare:69}
C.~Antony~R. Hoare.
\newblock An axiomatic basis for computer programming.
\newblock {\em Communications of the ACM}, 12(10):576--580, 1969.

\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}

### 2017-02-14 hoare/paper.tex
Add quadratic equation

@@
 
-For example lets take task on finding real roots for quadratic equation.
+For example lets take task on finding real roots for quadratic equation:
+$$ ax^2 + bx + c = 0$$
 
-We limits input set according to discriminant and set of real numbers in
-predicate P. Then we could find values by some program Q. And finally check
+We limits input domain as set of real numbers and according to discriminant in
+predicate P:
+$$ P \equiv (a, b, c \in \mathbb{R}) \wedge (b^2 - 4ac \geq 0)$$
+
+Then we could find values by some program Q. And finally check
 whether equation holds or not in R.

### 2017-02-14 hoare/paper.tex
Add postcondition for quadratic equation

@@
   Programming.
+  (WIP)
 
@@
 We limits input domain as set of real numbers and according to discriminant in
-predicate P:
+predicate P (suppose we used general quadratic formula):
 $$ P \equiv (a, b, c \in \mathbb{R}) \wedge (b^2 - 4ac \geq 0)$$
 
-Then we could find values by some program Q. And finally check
-whether equation holds or not in R.
+Then we could find values by some program Q.
+And finally check whether equation holds or not in R: 
+$$ R \equiv (X \subseteq \mathbb{R}) \wedge (\forall x \in X \bullet ax^2 + bx + c
+= 0)$$
 

### 2017-02-15 hoare/paper.tex
Write on tablet (about quadratic equation example and other)

@@
 So Hoare suggested triple \{P\} Q \{R\} \cite{hoare:69} in which developer need
-define 
-precondition (premises) and postcondition (conclusion) which hold for program Q
-after execution.
+define precondition (premises) and postcondition (conclusion) which hold for
+program Q after execution.
 
@@
 
-For next disscussion suppose that we define right P and R from some
+Or in other words, it helps us check that program do things right, but doesn't
+approve that what we want, is right things.
+
+For next discussion suppose that we define right P and R from some
 requirements. \\
@@
 
-Then we could find values by some program Q.
+Then we could find values by some program Q. Suppose as a result program returns
+set of values X (Don't think about implementation yet).
 And finally check whether equation holds or not in R: 
@@
 
-In the article Hoare states four basic rules:
+In the '69 article Hoare states four basic rules:
 \begin{itemize}
@@
 \end{itemize}
+But sometimes stated another two rules - skip axiom and condition rule. First
+one could be derived from axiom of assignment, and second from rule of
+iteration.\\
 
-Also it requires no side effects for Q (nothing changes except state which
-we define in pre and postconditions). Because in other case, we couldn't
-consider correctness of overall state. \\
+All this rules implicitly assumed the absence of side effects for Q (nothing
+changes except state which we define in pre and postconditions). Because in
+other case, we couldn't consider correctness of overall state. \\
 
@@
 statements (rule of composition D2), in which first part hold some postcondition
-R1 and second have R1 as precondition.
+R1 and second have R1 as precondition. But R1 could be arbitrary, for example
+$\not$ P.
 
-So nothing scary in that case if we still could prove correctness of our program.
+As we can see, there is nothing scary in that case if we still could prove
+correctness of our program.
+
+So now lets return to our example about real roots.
+
+Suppose we choose general quadratic formula for solving equation:
+
+$$ x_{1,2} = \frac{-b +- \square{b^2 - 4ac}}{2a}$$
+
+We know that $b^2 - 4ac$ not negative, but we don't know weather a is equals to
+zero. In this case equation became linear, but still have solutions.
+
+Lets try use substitution of x by this to prove that postcondition holds.
 

### 2017-02-15 hoare/paper.tex
Continue of equation example, awesome conclusions

--- 2017-02-14 hoare/paper.tex
@@
 
-So Hoare suggested triple \{P\} Q \{R\} \cite{hoare:69} in which developer need
-define 
-precondition (premises) and postcondition (conclusion) which hold for program Q
-after execution.
+So Hoare suggested triple $\{P\} Q \{R\}$ % \cite{hoare:69}
+in which developer need
+define precondition (premises) and postcondition (conclusion) which hold for
+program $Q$ after execution.
 
-This technique helps prove correctness of a program Q, according to defined
-logical predicates P and R.
+This technique helps prove correctness of a program $Q$, according to defined
+logical predicates $P$ and $R$.
 
@@
 provide ability to use common proof techniques for deducting correctness.
+In other words, it helps us check that program do things right, but doesn't
+affirm that what we want, is right things.
 
-For next disscussion suppose that we define right P and R from some
+For next discussion suppose that we define right $P$ and $R$ from some
 requirements. \\
@@
 Using axioms and inference rules what was proposed by Hoare, we could deduct
-that if P holds before execution of Q, and if Q terminates, then R will holds
-after execution (correctness of Q).\\
+that if $P$ holds before execution of $Q$, and if $Q$ terminates, then $R$ will
+holds after execution (correctness of $Q$).
 
-In 1969 paper Hoare didn't mentioned that Q should terminates, so he defined
+In 1969 paper Hoare didn't mentioned that $Q$ should terminates, so he defined
 rules just for deducting partial correctness (particular in rule D3). But later
-he add definition for total correctness too. \\
+he add definition for total correctness too.
 
@@
 
-We limits input domain as set of real numbers and according to discriminant in
-predicate P (suppose we used general quadratic formula):
+We restrict input domain as set of real numbers and according to discriminant in
+predicate $P$ (suppose we used general quadratic formula):
 $$ P \equiv (a, b, c \in \mathbb{R}) \wedge (b^2 - 4ac \geq 0)$$
 
-Then we could find values by some program Q.
+Then we could find values by some program $Q$. Suppose as a result program
+returns set of values $X$ (Don't think about implementation yet).
 And finally check whether equation holds or not in R: 
@@
 
+$P$ like a guard, filter input on what program will work correctly, and $R$
+check results to be correct.
+
 \section{Axioms and Rules}
 
-In the article Hoare states four basic rules:
+In the '69 article Hoare states four basic rules:
 \begin{itemize}
@@
 \end{itemize}
+But sometimes another two rules is stated - Skip Axiom and Condition Rule. First
+one could be derived from Axiom of Assignment, and second from Rule of
+Iteration.\\
 
-Also it requires no side effects for Q (nothing changes except state which
-we define in pre and postconditions). Because in other case, we couldn't
-consider correctness of overall state. \\
+All this rules implicitly assumed the absence of side effects for $Q$ (nothing
+changes except state which we define in pre and postconditions). Because in
+other case, we couldn't consider correctness of overall state. \\
 
 Using this rules we can make deductive proof from our premise to conclusion,
-that R will holds after execution.
+that $R$ will holds after execution.
 
@@
 
-It could be a case when program Q during execution
-change state which doesn't hold conditions P and R.
+It could be a case when program $Q$ during execution
+change state which doesn't hold conditions $P$ and $R$.
 
@@
 If we think about it in more details, we could decompose it to two sequence of
-statements (rule of composition D2), in which first part hold some postcondition
-R1 and second have R1 as precondition.
+statements (by rule of composition D2), in which first part hold some postcondition
+$R_1$ and second have $R_1$ as precondition. But $R_1$ could be arbitrary, for example
+$\neg P$.
 
-So nothing scary in that case if we still could prove correctness of our program.
+As we can see, there is nothing scary in this case if we still could prove
+correctness of our program. \\
 
-P like a guard, filter input on what program will work correctly, and R check
-results to be correct.
+Lets return to our example about real roots.
+
+Suppose we choose general quadratic formula for solving equation:
+$$ x_{1,2} = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$$
+
+We know that $b^2 - 4ac$ not negative, but we
+don't know weather $a$ is equals to zero.
+In this case equation became linear, but still have solutions.
+
+Suppose we didn't see that general formula is derived from equation. Lets just
+apply substitution all occurrence of $x$ in postcondition by stated expressions.
+
+After polynomial expansion we could see that equation takes the form:
+$$\frac{b^2}{2a} \mp \frac{b\sqrt{b^2 - 4ac}}{2a} - c - \frac{b^2}{2a} \pm
+\frac{b\sqrt{b^2 - 4ac}}{2a} + c = 0$$
+
+I little bit cheated, because I already reduce some parts by $a$, but
+nevertheless, after reducing the remaining we get:
+$$ 0 = 0 $$
+
+But we should remember, that we could reduce only by not zero value, thus
+postcondition holds only if $a \neq 0$.
+
+Lets predicate $B \equiv a \neq 0$, then precondition for our statement:
+$$ P_1 \equiv P \wedge B $$
+
+By the manipulations above we actually proved that our statement correct, with
+respect to precondition $P_1$ and postcondition $R$.
+
+Actually we could replace whole program with one assignment:
+$$ X = \emptyset $$
+
+And it will be correct program, according to specification. That's why I
+stressed that it doesn't help obtain right predicates.
 
@@
 
-However precondition and postconditions could be almost same. This looks ugly.
+% However precondition and postconditions could be almost same. This looks ugly.
 
-firstly we should define the
-truth.
+% firstly we should define the
+% truth.
 
-So directed tests could show that implementation isn't work properly. This is
-helpful during development and evolving software product.
+% So directed tests could show that implementation isn't work properly. This is
+% helpful during development and evolving software product.
 

### 2017-02-16 hoare/paper.tex
Desribe rule of consequence

@@
 
+And we could notice that predicate $P_1$ is stronger that predicate $P$. By
+``stronger'' we mean that $P_1 \implies P$, or informally, if predicate $P_1$
+holds for some values, predicate $P$ will holds either. But opposite is not true
+(in our case $x = 0$ doesn't hold in $P_1$).
+
+So we could make preconditions stronger, program will still be correct.
+
+With postcondition predicates, on the contrary, we couldn't make it
+stronger.
+
+Because if we don't change program $Q$, it will produce range of
+results, which satisfy weaker predicate.
+
+This range is superset of set which satisfy stronger predicate.
+
+So $Q$ will produce values which don't satisfy postcondition - program will be
+incorrect according to stronger predicate.
+
+But we could make postcondition weaker, program will still be correct.
+
+This is actually Rule of Consequence stated in Hoare paper.
+
+But lets return to our example.
+
+% We prove correctness of statement on conditions $P_!$ and $Q$. But we should
+% consider
+
+In case when $x = 0$ (or $P \wedge \neg B$) equation takes the form:
+$$ bx + c = 0 \text{, and } x = \frac{-c}{b}$$
+
+Yet again this remains true when $C \equiv b \neq 0$ is true, so this statement
+is correct according to precondition $P \wedge \neg B \wedge C$ and
+postcondition $R$.
+
+As you can see, we try discharge proof by cases, and then, using Condition
+Rule (or Rule of Iteration), prove correctness of their combination.
+
+In case when $b = 0$ or $\neg C$, equation takes the form:
+$$ c = 0$$
+
+Either $c = 0$ and equation holds for all $x \in \mathbb{R}$, or $c \neq 0$ and
+equation
+
 Actually we could replace whole program with one assignment:

### 2017-02-16 hoare/paper.tex
Seems done with equation example

@@
 % \fancyfoot[C]{\thepage} % Custom footer text
+
+\usepackage{amsmath, amsfonts}
+\renewcommand{\implies}{\Rightarrow}
 
@@
 
+%------------------------------------------------
+
 \section{Purpose and Applications}
@@
 
+%------------------------------------------------
+
 \section{Axioms and Rules}
@@
 Suppose we choose general quadratic formula for solving equation:
-$$ x_{1,2} = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$$
+\begin{equation}
+ x_{1,2} = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
+\end{equation}
 
@@
 Suppose we didn't see that general formula is derived from equation. Lets just
-apply substitution all occurrence of $x$ in postcondition by stated expressions.
+apply substitution all occurrence of $x$ in postcondition by stated expression (1).
 
@@
 
-Lets predicate $B \equiv a \neq 0$, then precondition for our statement:
-$$ P_1 \equiv P \wedge B $$
+Lets predicate $A \equiv a \neq 0$, then precondition for our statement:
+$$ P_1 \equiv P \wedge A $$
 
@@
 stronger.
-
 Because if we don't change program $Q$, it will produce range of
 results, which satisfy weaker predicate.
-
 This range is superset of set which satisfy stronger predicate.
@@
 
-This is actually Rule of Consequence stated in Hoare paper.
+This is actually Rule of Consequence stated in Hoare paper.\\
 
@@
 
-In case when $x = 0$ (or $P \wedge \neg B$) equation takes the form:
-$$ bx + c = 0 \text{, and } x = \frac{-c}{b}$$
+In case when $x = 0$ (or $P \wedge \neg A$) equation takes the form:
+\begin{equation}
+ bx + c = 0 \text{, and } x = \frac{-c}{b}
+\end{equation}
 
-Yet again this remains true when $C \equiv b \neq 0$ is true, so this statement
-is correct according to precondition $P \wedge \neg B \wedge C$ and
+Yet again, this remains true when $B \equiv b \neq 0$ is true, so this statement
+is correct according to precondition $P \wedge \neg A \wedge B$ and
 postcondition $R$.
@@
 
-In case when $b = 0$ or $\neg C$, equation takes the form:
+In case when $b = 0$ or $\neg B$, equation takes the form:
 $$ c = 0$$
 
-Either $c = 0$ and equation holds for all $x \in \mathbb{R}$, or $c \neq 0$ and
-equation
+And again, two cases. Either $c = 0$ and equation holds for all
+$x \in \mathbb{R}$:
+\begin{equation}
+  X = \mathbb{R}
+\end{equation}
 
-Actually we could replace whole program with one assignment:
+Or $c \neq 0$ and equation doesn't have solutions:
+\begin{equation}
+  X = \emptyset
+\end{equation}
+
+So lets combine this four statements and prove correctness of our program.
+
+We will use Condition Rule, which could be derived from Rule of Iteration:
+$$\frac{\{P \wedge A\} Q_1 \{R\}, \{P \wedge \neg A\} Q \{R\}}
+{\{P\} \textbf{ if } A \textbf{ then } Q_1 \textbf{ else } Q \textbf{ end } \{R\}}$$
+
+$Q_1$ correspond to equation (1), $Q_2$ to (2) and so on. $Q$ represent
+nested statements:
+
+Therefore, we prove that program holds postcondition $R$ for all values
+satisfying $P$ and it is total correct:
+\begin{align*}
+  \{P\} & \textbf{ if } A \textbf{ then } Q_1 \textbf{ else } \\
+        & \hspace{1em} \textbf{ if } B \textbf{ then } Q_2 \\
+        & \hspace{2em} \textbf{ if } C \textbf{ then } Q_3 \textbf { else } Q_4 \textbf { end } \\
+        & \hspace{1em} \textbf{ end }\\
+        & \textbf{ end } \{R\}
+\end{align*}
+
+\rightline{{\rm $\Box$}}
+
+%------------------------------------------------
+
+Actually we could replace whole program with just one assignment:
 $$ X = \emptyset $$
 
-And it will be correct program, according to specification. That's why I
-stressed that it doesn't help obtain right predicates.
+And prove it correctness according to $P$ and $R$ too, because universal
+quantification expressed that every member of domain is satisfied some
+predicate. This is true, when there is no members.
+
+That's why I stressed that it doesn't help obtain right predicates.
 

### 2017-02-16 hoare/paper.tex
Shortcomings and conclusion

@@
 
-  This is review on C.A.R. Hoare's paper, An Axiomatic Basis for Computer
-  Programming.
-  (WIP)
+  This paper is review on C.A.R. Hoare's paper, An Axiomatic Basis for Computer
+  Programming \cite{hoare:69}.
+
+  Basically an attempt is made to apply technique for reasoning about fictional
+  example.
 
@@
 ``People are not perfect'', this is what we learned so fourth. 
-Every one could make a mistake. In some areas it could be unnoticed, but in
+Everyone could make a mistake. In some areas it could be unnoticed, but in
 others it could cause serious consequences.
@@
 failures cost increases, we will pay for diligence and accuracy as much as
-needed.
+needed.\\
 
@@
 program $Q$ after execution.
+
+Also he state axioms and rules to make inference about statements.
 
@@
 rules just for deducting partial correctness (particular in rule D3). But later
-he add definition for total correctness too.
+he add definition for total correctness too. \\
 
@@
 We restrict input domain as set of real numbers and according to discriminant in
-predicate $P$ (suppose we used general quadratic formula):
+predicate $P$ (suppose we used general quadratic formula and restrict domain by
+cases in which equation has real roots):
 $$ P \equiv (a, b, c \in \mathbb{R}) \wedge (b^2 - 4ac \geq 0)$$
@@
 Then we could find values by some program $Q$. Suppose as a result program
-returns set of values $X$ (Don't think about implementation yet).
+returns set of values $X$.
 And finally check whether equation holds or not in R: 
@@
 changes except state which we define in pre and postconditions). Because in
-other case, we couldn't consider correctness of overall state. \\
+other case, we couldn't consider correctness of overall state.
 
@@
 \begin{equation}
- x_{1,2} = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
+ X = \Big\{ \frac{-b \pm \sqrt{b^2 - 4ac}}{2a} \Big\}
 \end{equation}
@@
 
-Suppose we didn't see that general formula is derived from equation. Lets just
-apply substitution all occurrence of $x$ in postcondition by stated expression (1).
+Suppose we didn't see that general formula is directly derived from equation.
+Lets just apply substitution all occurrence of $x$ in postcondition by stated
+expression (1). \\
 
@@
 By the manipulations above we actually proved that our statement correct, with
-respect to precondition $P_1$ and postcondition $R$.
+respect to precondition $P_1$ and postcondition $R$. \\
 
-And we could notice that predicate $P_1$ is stronger that predicate $P$. By
+We could notice that predicate $P_1$ is stronger that predicate $P$. By
 ``stronger'' we mean that $P_1 \implies P$, or informally, if predicate $P_1$
@@
 incorrect according to stronger predicate.
-
 But we could make postcondition weaker, program will still be correct.
 
-This is actually Rule of Consequence stated in Hoare paper.\\
+This is Rule of Consequence, stated in Hoare paper.\\
 
@@
 \begin{equation}
- bx + c = 0 \text{, and } x = \frac{-c}{b}
+  bx + c = 0 \text{, and set } X = \Big\{ \frac{-c}{b} \Big\}
 \end{equation}
@@
 is correct according to precondition $P \wedge \neg A \wedge B$ and
-postcondition $R$.
+postcondition $R$. \\
 
-As you can see, we try discharge proof by cases, and then, using Condition
+As you can see, we try to discharge proof by cases, and then, using Condition
 Rule (or Rule of Iteration), prove correctness of their combination.
@@
 $x \in \mathbb{R}$:
+
 \begin{equation}
@@
 
-So lets combine this four statements and prove correctness of our program.
+So lets combine this four statements and prove correctness of our program. \\
 
@@
 $Q_1$ correspond to equation (1), $Q_2$ to (2) and so on. $Q$ represent
-nested statements:
+nested statements, including control structures and statements $Q_2, Q_3, Q_4$.
+
+In the same way we prove that $Q$ is correct.\\
 
@@
 \rightline{{\rm $\Box$}}
+\vspace{1em}
 
-%------------------------------------------------
+That's it. With one significant remark.
 
-Actually we could replace whole program with just one assignment:
+Actually we could replace whole program
+by just one assignment: 
 $$ X = \emptyset $$
@@
 quantification expressed that every member of domain is satisfied some
-predicate. This is true, when there is no members.
+predicate. This is true, when there is no members. \\
 
-That's why I stressed that it doesn't help obtain right predicates.
-
-% But sometimes it could be not so straightforward. Lets return to our example
-% about real roots. Suppose quadratic coefficient a could be equals 0.
+\textit{That's why I stressed, that it doesn't help obtain right predicates. }
+%------------------------------------------------
 
@@
 % There is set of rules (or axioms) which we could use to deduct correctness of a program.
-
-% However precondition and postconditions could be almost same. This looks ugly.
-
-% firstly we should define the
-% truth.
 
@@
 
+As we can see with equation example, deriving good pre and postconditions is not
+trivial task. It could be ambiguous when done wrong.
+
+The first thing that comes to mind change postcondition:
+$$ R^\prime \equiv R \wedge \neg(\exists x \not\in X \bullet ax^2 + bx + c = 0)$$
+
+To check that program returns all solutions to equation. But, in practice it
+very difficult prove existential quantification's.
+
+Also quantifiers cannot be expressed as boolean expression in ordinary
+languages, which doesn't support it \cite{meyer:90}.
+
+So defining predicates is weak part of method overall.
+
+% But as soon as we stated right one, we could gain confidence in reliability and
+% correctness of our software, as much as we believe in deductive logic.
+
+%------------------------------------------------
+
 \section{Conclusion}
+
+Verification software in general is difficult task.
+
+With constantly increases complexity of software and it dependencies is hard to
+tell certainly, whether particular program will accomplish task right or not.
+
+This proof technique not only had a great influence on evolution of software
+verification, but also alleviate some of the other problems which exist in
+computing word.
+
+Strict definition of subroutines' variables before and after execution, helps
+understand purpose and accomplishment for potential user.
+
+Lets not forget that Halting problem is undecidable over Turing machines.
+
+With technique which Hoare suggests we can prove total correctness of program.
+
+But after all, we can gain confidence in reliability and correctness of our
+software as much as we believe in deductive logic, validity of our
+assumptions and reliability of the electronics.
 
@@
 
+\bibitem[Mey90]{meyer:90}
+Bertrand Meyer.
+\newblock Introduction to the Theory of Programming Languages.
+\newblock {\em Prentice-Hall, Inc.}, 0-13-498510-9, 1990.
+
 \end{thebibliography}

### 2017-02-17 hoare/paper.tex
Finish conclusion for the first version

@@
 Also quantifiers cannot be expressed as boolean expression in ordinary
-languages, which doesn't support it \cite{meyer:90}.
+languages, which doesn't support it \cite{meyer:90}.\\
 
-So defining predicates is weak part of method overall.
+Another disadvantage of software verification and modeling overall is a cost.
+Because it requires much more precision and rigor, they are always going to be
+more expensive than traditional approaches to engineering.
 
@@
 
+Nowadays nobody writes code from scratch. We rely on quality of underlying
+layers. But we can't prove correctness for our program, until we prove
+correctness for all underlying layers. \\
+
+It could be the case, when we don't want to spent huge amount of time to prove
+software, but not in critical industries. The fundamental basis is required in
+these areas.
+
+Like with the bricks, we need to construct base for next level of abstraction,
+proving that we don't make mistake which could cost lives.
+Can be used different techniques of proof. But most important thing is to be
+confident in theory and in practice.
+\\
+
 This proof technique not only had a great influence on evolution of software
-verification, but also alleviate some of the other problems which exist in
-computing word.
+verification, but also alleviate some of the other difficulties which appears in 
+computing word.\\
 
-Strict definition of subroutines' variables before and after execution, helps
-understand purpose and accomplishment for potential user.
+One of them is a documentation of source code.
+Strict definition of subroutine's variables before and after execution, helps
+potential user understand purpose and how use it.
 
-Lets not forget that Halting problem is undecidable over Turing machines.
+It might be a better documentation than the description written at four pages.
 
-With technique which Hoare suggests we can prove total correctness of program.
+Because in principle we don't need understand how to work black box inside, if
+we sure how we can use it outside.
+Understanding will come with time.
+\\
 
-But after all, we can gain confidence in reliability and correctness of our
-software as much as we believe in deductive logic, validity of our
-assumptions and reliability of the electronics.
+Another issue is maintenance existing code base.
+First of all, we won't receive messages about errors, will not spent time on
+fixing and debugging.
+
+We will spend time on modeling and proofs instead.
+\\
+
+Slowly but surely, after all, we will gain confidence in reliability and
+correctness of our software as much as we believe in deductive logic, validity
+of our assumptions and reliability of the electronics.
 
